\SW: II. New Gravitational Lens Candidates from the CFHTLS Discovered through Citizen Science

### 2015-05-05 doc/sw-cfhtls.tex
corrections from co-authors v1

The last few decades have seen a rise in the discoveries of strong
gravitational lenses owing to the plethora of interesting applications
lenses have in astrophysics and cosmology. Strong lenses are routinely
used to probe the dark matter distribution from galaxy
\citep[e.g.][]{Koopmans2006,Barnabe2009,Leier2011,Sonnenfeld2015} to groups-cluster
scales
\citep[e.g.][]{Limousin2008,Zitrin2011,Oguri2012,More2012,Newman2013}, to
study distant young galaxies by using the lensing magnification as a
natural telescope \citep[e.g.][]{Zitrin2009,Zheng2012,Whitaker2014}, to
test the cosmological model by constraining cosmological parameters such
as the Hubble constant and dark energy
\citep[e.g.][]{Suyu2010,Collett2012,Collett2014,Sereno2014} and many more. Strong
lenses are rare since a foreground massive object needs to be
sufficiently aligned with a distant background source to produce
multiple images. Nevertheless, systematic lens searches have led to the
discovery of over 500 lenses till
date\footnote{http://admin.masterlens.org/index.php}.

@@

\GZ was the first citizen science project in astronomy, and addressed
the problem of how to  classify large numbers of galaxies by their
morphology \citep{Lintott2008}.  From the early results based on
morphological classifications of galaxies, to several new unexpected and
interesting discoveries, such as that of green pea galaxies
\citep{Cardamone2009,Jaskot2013} and Hanny's Voorwerp
\citep{Lintott2009,Keel2012}, \GZ has been able to harness the potential
of citizen scientists.  Since then astronomy and non-astronomy projects
have been launched under the citizen science web portal Zooniverse
(http://zooniverse.org). The task of finding gravitational lenses is
significantly challenging given that the lens systems show such
complexity and that they are rare. To add to the challenge, not many
citizen scientists are expected to be aware of the phenomenon of
gravitational lensing, and the resulting image configurations.  With
these significant challenges at hand, we designed the \sw project to
enable the discovery of lenses through citizen science.  In a companion
paper, Paper I (Marshall et al., in prep.), we describe the design of
\sw and how the entire system functions as a discovery service.  In this
paper (Paper II hereafter), we describe our first lens search using data
from the Canada-France-Hawaii Telescope Legacy Survey
(\cfhtls\footnote{\texttt{http://www.cfht.hawaii.edu/Science/CFHTLS/}}).
{\bf In \citet{Kung2015}, we describe the design of a collaborative
mass modelling tool that can be used by citizen scientists
.}

@@

The \cfhtls is a photometric survey in five optical bands
($u^*g'r'i'z'$) carried out with the wide-field imager MegaPrime which
has a 1~deg$^2$ field-of-view and a pixel size of 0.186\arcsec
\citep{Gwyn2012}. The \cfhtls \textsc{WIDE} covers a total
non-overlapping area of 160~deg$^2$ on the sky and consists of four
fields W1, W2, W3 and W4. The field W1 has the largest sky coverage of
63.7~deg$^2$. The fields W2 and W4 have similar sky coverages of
22.6~deg$^2$ and 23.3~deg$^2$, respectively\footnote{These numbers are
estimated from http://terapix.iap.fr/cplt/T0007/doc/T0007-doc.pdf}.  The
field W3 has a sky coverage of 44.2~deg$^2$ and is more than twice as
large as W2 and W4.

The \cfhtls imaging is very homogeneous and has good image quality. Most
of the lensed arcs are much brighter in the $g$ band thus, deep imaging
in this band is desirable. The limiting magnitude is 25.47 for the $g$
band which goes the deepest among all of the five bands. The mean seeing
in the $g$ band is 0.78\arcsec. The zero point to convert flux to AB
magnitude for all bands is 30. These characteristics make \cfhtls ideal
to do visual inspection for finding lenses.  We use the stacked images
from the final T0007 release taken from the Terapix
website\footnote{{http://terapix.iap.fr/cplt/T0007/doc/T0007-doc.pdf}} for this
work.

@@

In the thesis dissertation of \citet{Thanjavur2009}, 9 promising and 2
low probability candidates are reported which were discovered
serendipitiously. These detections were made during the visual
inspection of the \cfhtls images as part of data reduction procedures
for the Weak Lensing survey \citep{Benjamin2007}.

@@

The work by \citet{Maturi2014} used the arc finding code of
\citet{Seidel2007} and color properties of typical arcs to optimize arc
detection. This new approach was tested on the CFHTLS-Archive-Research
Survey \citep[CARS,][]{Erben2009} which covers an area of 37~sq.~deg
only and this entire data was also visually inspected by the authors to
estimate completeness and purity of their robotic search. They found 29
candidates with the robotic search alone and 41 candidates through pure
visual inspection some of which are known from previous searches. Most
of these candidates are medium-low
probability\footnote{http://www.ita.uni-heidelberg.de/$\sim$maturi/Public/arcs}.

The \rf and the \af searches are the only searches that make use of a
lens finding algorithm and that has been run on the entire \cfhtls
imaging.  Thus, these can be considered as our reference sample of known
lenses from robotic searches. For the purposes of transparency and to
help with the training, the volunteers participating in \sw-\cfhtls lens
search were made aware of these two known lens samples. Images
containing the systems from the \rf and the \af samples were labelled as
``Known Lens Candidates'' in the \sw discussion forum,
TALK\footnote{http://talk.spacewarps.org/}, where volunteers have the
opportunity to discuss their findings with fellow volunteers and the
science team. In this paper, we refer to the sample of 330 \rf and 108
\af lens candidates as the sample of ``known lens candidates'' and the
sample of confirmed (or most promising) 33 \rf and 26 \af as the sample
of ``known lenses''. Note that the ``known lens'' sample is a subset of
the ``known lens candidates'' sample.

@@

\subsection{Image Presentation in \sw}
\label{sec:data:impres}
In order to perform a blind lens search over the entire \cfhtls
\textsc{wide}, we present the volunteers with cutouts of images selected
from the survey region. We briefly describe the image presentation here
for completeness but more information can be found in Paper I. We use
the $g$, $r$ and $i$-band imaging from \cfhtls which are most useful for
visual identification of lenses.  We made color composite images using
the publicly-available code, \humvi\footnote{The open source  color image
composition code used in this work is available from
\texttt{http://github.com/drphilmarshall/HumVI}} following the
prescription of \citet{Lupton2004}. The color scales are chosen to
maximize the contrast between faint extended objects. These parameters
are then fixed during the production of all the tiles, in order to allow
straightforward comparison between one image and another, and for
intuition to be built up about the appearance of stars and galaxies
across the survey.

We extract contiguous cutouts of size 81.84\arcsec\ (440 pixels) with an
overlapping region of 10\arcsec (54 pixels) between the neighbouring
cutouts. This  results in a catalog of $\sim$ 430 000 cutouts for the entire \cfhtls
\textsc{wide} region. The size of the individual cutout is determined by
optimising factors such as the typical angular scales of gravitational
lenses, the number of objects seen in a single cutout and the total
number of image cutouts in the survey. If a lens
candidate happens to be too close to the edge of a cutout, then the overlap
between neighbouring cutout allows a volunteer to get a clearer view
of the same candidate in at least one of the cutouts. We note that since
the images are shown randomly, a volunteer may not necessarily come
across the neighbouring cutout unless {\bf they classify} a large
number of images. This is not a problem since our user base is extremely
large and we receive multiple classifications for the same cutout.

@@

The simulated lenses are important to train citizen scientists who are
novice to the task of finding lenses but they are also crucial for
analyzing the classifications performed by the citizen scientists (more
details can be found in Paper I but see \Sref{sec:swap} for a brief
summary). In this section, we describe the framework used for generating the
simulated lens sample and give details of the sample itself along with
some known limitations.

@@

For the purpose of generating simulated lens systems, we divide them
in to two main categories a) galaxy scale lenses b) and group or
cluster-scale lenses. We further subdivide galaxy scale lenses based on the
type of the background sources, namely galaxies and quasars. We do not simulate
group scale - quasar lenses as they are expected to be even more rare. We now
describe our procedure to generate these different types of lens systems.

@@

Since we want to produce realistic looking lens systems, we simulate
lenses in each of the five \cfhtls~filters. The colors of the background
galaxies are drawn randomly from the photometric CFHTLenS catalog
\citep{Hildebrandt2012,Erben2013}.  Similarly, we use a quasar catalog
from the SDSS Data Release 9 \citep{Paris2012} from which colors are
drawn to simulate quasar lenses. Next, we assumed a Gaussian
profile\footnote{This was due to an oversight. We intend to use either
an exponential or {\bf de Vaucouleurs'} profile.} for
the galaxies. The ellipticity and the position angle (PA) are drawn
randomly from within the range given in Table~\ref{tab:thresh}. The
effective radius of the galaxy is estimated from the Luminosity$-$size
relation \citep{Bernardi2003} (with a redshift scaling, to account for
size evolution) given by
\be
R_{\rm eff}= 10^{0.52} \frac{L_r^{2/3}}{{(1+z_s)}^2} \, {\bf Kpc}
\ee
where $L_r=L_s/10^{10.2} {\bf L_\odot}$. On the other hand, quasars are assumed to be
point sources and the PSF, with which quasars are convolved, is assumed
to have a Gaussian profile. The full-width-at-half-maximum of PSF is
equated to that of the median seeing for every filter. The median seeing
values are taken from Table 4 of the official Terapix T0007 release
explanatory document \footnote{
http://terapix.iap.fr/cplt/T0007/doc/T0007-doc.pdf}.

@@

\subsection{Simulated Lens Sample and Catalog Description}

@@

We produce catalogs with lens and source properties for each of the
three types of lenses.  The catalogs typically have lens position,
redshift, magnitudes, Einstein radius, ellipticity (whenever available)
and shear (for galaxy-scale lenses only). For the background sources, we
provide the offset from the lens center, redshift, magnitudes, total
magnification, number of lensed images. Additionally, when possible,
ellipticity and effective radius of the background galaxies have also
been provided. These catalogs are available from
https://github.com/anupreeta27/SIMCT and simulated lens sample can be
requested from the authors.

@@

\subsection{Limitations of the simulated sample}
The simulated lens sample, although realistic, is not perfect, due to
the simplicity of the lensing models and our limited understanding of
the uncertainties in the model parameters. Comments from citizen
scientists were very helpful in order to sort some of these failures.
Here, we describe some of the cases or aspects in which the simulations
were known to have failed to look realistic. The fraction of such
simulated sample is small (roughly 5 percent)\footnote{This estimate is
based on the \#simfail from TALK, the discussion forum.}.

The parameters required by various scaling relations and the models
primarily depend on the photometry of the galaxies, groups and quasars
detected in the survey. For galaxy-scale lenses, the lensing galaxies at
higher redshifts or which are fainter have poor photometric redshift
measurements. Consequently these galaxies are assigned a wrong
luminosity and velocity dispersion estimates. This then results in
simulated lenses which look implausible or unrealistic.  For example,
the lensed images for some of the failed simulations have larger image
separation than what you expect from the luminosity and/or size of the
galaxy. We roughly expect mass to follow light so more massive galaxy
typically looks brighter and/or bigger.

@@

We selected a sample of 450 duds for the \StageOne classification in \sw
and a sample of 500 false positives for the \StageTwo inspection. The
sample of false positives was selected from the candidates which passed
the \StageOne of \sw. We note that this is the first time, we have a
systematically compiled sample of visually inspected false positives by
the \sw volunteers and categorized by the science team. We produce
additional larger sample of a few thousand FPs by scanning through the
low probability images after the completion of \StageTwo. All of these
data products will be made available at
http://spacewarps.org/\#/projects/CFHTLS/. Such a sample has tremendous
utility for training and testing of various lens finding algorithms
\citep[e.g.,][]{Chan2014}.

\begin{figure}
\begin{center}
\includegraphics[scale=0.6]{cand_fp_P_frac.pdf}
\caption{ \label{fig:stackP}
Distribution of different types of candidates as a function of the posterior
probability P. The types of the candidates are the FPs, the new candidates and
the known candidates. The new and known candidates have higher detection rate
for higher values of P, as expected.}
\end{center}
\end{figure}

@@

\section{Methodology to produce the \sw-\cfhtls Lens Sample}
\label{sec:swap}
The \sw works as a single unified system which uses the method of visual
inspection to find gravitational lenses. For the first \sw lens search,
the volunteers were shown images at two stages. At \StageOne, volunteers
are requested to undertake a rapid inspection to select lens candiates
ranging from possible lenses to almost certain lenses. At \StageTwo,
volunteers carefully inspect the candidates from \StageOne and select
only promising lens candidates. A daily snapshot of classifications
performed by volunteers are provided to the science team every night.
This daily batch is analyzed by the Space Warps Analysis Pipeline
(SWAP). The philosophy and the details of SWAP are described in detail
in Paper I, here we briefly summarize how it works.

@@

\subsection{\sw-\cfhtls Lens Sample}
\label{sec:swlens}

In this section, we describe the \sw-\cfhtls lens sample (henceforth,
called the \sw lens sample). We find a total of 141 candidates with
G$\ge1.3$ (medium-high probability sample) out of which {\bf 59} are new
systems. This sample is further divided as follows. We have a total of
50 candidates with $1.3\le$G$<2$ (medium probability sample) out of
which 30 are new. The quality of candidates in this category is such
that at least one of the inspectors thought the candidate is most
probably a lens (that is, a grade of 2) and a second inspector thought
this is possibly a lens (that is, a grade of 1). Among our high
probability sample (G$\ge$2), there are a total of 91 candidates out of
which {\bf 29} are new. In this category, the minimum grade by all of the
inspectors was 2 suggesting that the candidates are probably or almost
certainly lenses according to all of the inspectors. To avoid
duplication, further information on \sw candidates that were previously
identified in the literature will be made available at
http://spacewarps.org/\#/projects/CFHTLS/ whereas the sample of new lens
candidates with G$\ge 1.3$ (medium-high probability) is presented in
this paper (see \Sref{sec:results:newcand}).

We find a total of 288 (and {\bf 245} new) candidates with averaged grade
$0<$G$<1.3$ (low probability sample) which means - in the worst case, at
least one of the inspector thought the candidate is possibly a lens and
in the best case, all three inspectors thought the candidate is possibly
a lens (that is, a grade of 1)\footnote{If grades from the inspectors
were found to be discrepant by 2 or more, these were discussed and
re-graded to resolve the discrepancy.}. Further information on the low
probability sample such as their positions and images will be available
at http://spacewarps.org/\#/projects/CFHTLS/. Note that if all of the
inspectors gave a grade of 0 to a candidate, then it is discarded from
the sample.

@@

We show images of our new sample in \Fref{fig:lc} which is arranged
first in the decreasing order of their grades and then increasing order
of their positions. As the first lens search was a blind search with no
preselection of candidates from any algorithm, we find various types of
lenses, as expected from such a search. The final sample consists of
both galaxy and groups-scale lens candidates. There are detections of
elongated arcs and some interesting point-like quasar lensed images.
Most of them are brighter in bluer $g$ band but some candidates brighter
in the redder $i$ band are also found. Since robotic lens searches often
look for blue colored lensed features, they are very much likely to miss
such interesting lens candidates. We did not find any examples of
exotic lens candidates from the visually inspected P$>0.3$ sample. There
may be some interesting candidates lurking in TALK but these will have
to be sifted through

@@

The power of citizen scientists also lies in the high dynamic range that
allows us to find systems with very from short (thick) to long (thin)
arcs, from highly compact to low surface brightness images, from round
and point-like to elongated and curved images, from blue to red, from
regular to exotic kinds of lenses while keeping the false positive rate
low compared to algorithms. Discovery of this large sample of completely
new candidates missed by some of these algorithms demonstrates that the
\sw system is functioning well and the self-taught citizen scientists
have managed to beat the robots in some sense.

>>>>

The last few decades have seen a rise in the discoveries of strong
gravitational lenses owing to the plethora of interesting applications
lenses have in astrophysics and cosmology. Strong lenses are routinely
used to probe the dark matter distribution from galaxy
\citep[e.g.][]{Koopmans2006,Barnabe2009,Leier2011,Sonnenfeld2015} to groups-cluster
scales
\citep[e.g.][]{Limousin2008,Zitrin2011,Oguri2012,More2012,Newman2013}, to
study distant young galaxies by using the lensing magnification as a
natural telescope \citep[e.g.][]{Zitrin2009,Zheng2012,Whitaker2014}, to
test the cosmological model by constraining cosmological parameters such
as the Hubble constant and dark energy
\citep[e.g.][]{Suyu2010,Collett2012,Collett2014,Sereno2014} and many more. Strong
lenses are rare since a foreground massive object needs to be
sufficiently aligned with a distant background source to produce
multiple images. Nevertheless, systematic lens searches have led to the
discovery of over 500 lenses till
date\footnote{\texttt{http://admin.masterlens.org/index.php}}.

@@

\GZ was the first citizen science project in astronomy, and addressed
the problem of how to  classify large numbers of galaxies by their
morphology \citep{Lintott2008}.  From the early results based on
morphological classifications of galaxies, to several new unexpected and
interesting discoveries, such as that of green pea galaxies
\citep{Cardamone2009,Jaskot2013} and Hanny's Voorwerp
\citep{Lintott2009,Keel2012}, \GZ has been able to harness the potential
of citizen scientists.  Since then astronomy and non-astronomy projects
have been launched under the citizen science web portal Zooniverse
(\texttt{http://zooniverse.org}). The task of finding gravitational lenses is
significantly challenging given that the lens systems show such
complexity and that they are rare. To add to the challenge, not many
citizen scientists are expected to be aware of the phenomenon of
gravitational lensing, and the resulting image configurations.  With
these significant challenges at hand, we designed the \sw project to
enable the discovery of lenses through citizen science.  In a companion
paper, Paper I (Marshall et al., in prep.), we describe the design of
\sw and how the entire system functions as a discovery service.  In this
paper (Paper II hereafter), we describe our first lens search using data
from the Canada-France-Hawaii Telescope Legacy Survey
(\cfhtls\footnote{\texttt{http://www.cfht.hawaii.edu/Science/CFHTLS/}}).
{\bf In \citet{Kung2015}, we describe the design of a collaborative
mass modelling tool that can be used by citizen scientists
.}

@@

The \cfhtls is a photometric survey in five optical bands
($u^*g'r'i'z'$) carried out with the wide-field imager MegaPrime which
has a 1~deg$^2$ field-of-view and a pixel size of 0.186\arcsec
\citep{Gwyn2012}. The \cfhtls WIDE covers a total
non-overlapping area of 160~deg$^2$ on the sky and consists of four
fields W1, W2, W3 and W4. The field W1 has the largest sky coverage of
63.7~deg$^2$. The fields W2 and W4 have similar sky coverages of
22.6~deg$^2$ and 23.3~deg$^2$, respectively\footnote{These numbers are
estimated from
\texttt{http://terapix.iap.fr/cplt/T0007/doc/T0007-doc.pdf}}.  The field
W3 has a sky coverage of 44.2~deg$^2$ and is more than twice as large as
W2 and W4.

The \cfhtls imaging is very homogeneous and has good image quality. Most
of the lensed arcs are much brighter in the $g$ band thus, deep imaging
in this band is desirable. The limiting magnitude is 25.47 for the $g$
band which goes the deepest among all of the five bands. The mean seeing
in the $g$ band is 0.78\arcsec. The zero point to convert flux to AB
magnitude for all bands is 30. These characteristics make \cfhtls ideal
to do visual inspection for finding lenses.  We use the stacked images
from the final T0007 release taken from the Terapix
website\footnote{\texttt{http://terapix.iap.fr/cplt/T0007/doc/T0007-doc.pdf}} for this
work.

@@

In the thesis dissertation of \citet{Thanjavur2009}, 9 promising and 2
low probability candidates are reported which were discovered
serendipitously. These detections were made during the visual
inspection of the \cfhtls images as part of data reduction procedures
for the Weak Lensing survey \citep{Benjamin2007}.

@@

The work by \citet{Maturi2014} used the arc finding code of
\citet{Seidel2007} and color properties of typical arcs to optimize arc
detection. This new approach was tested on the CFHTLS-Archive-Research
Survey \citep[CARS,][]{Erben2009} which covers an area of 37~sq.~deg
only and this entire data was also visually inspected by the authors to
estimate completeness and purity of their robotic search. They found 29
candidates with the robotic search alone and 41 candidates through pure
visual inspection some of which are known from previous searches. Most
of these candidates are medium-low
probability\footnote{\texttt{http://www.ita.uni-heidelberg.de/$\sim$maturi/Public/arcs}}.

The \rf and the \af searches are the only searches that make use of a
lens finding algorithm and that has been run on the entire \cfhtls
imaging.  Thus, these can be considered as our reference sample of known
lenses from robotic searches. For the purposes of transparency and to
help with the training, the volunteers participating in \sw-\cfhtls lens
search were made aware of these two known lens samples. Images
containing the systems from the \rf and the \af samples were labelled as
``known lens candidates'' in the \sw discussion forum,
TALK\footnote{\texttt{http://talk.spacewarps.org/}}, where volunteers have the
opportunity to discuss their findings with fellow volunteers and the
science team. In this paper, we refer to the sample of 330 \rf and 108
\af lens candidates as the sample of ``known lens candidates'' and the
sample of confirmed (or most promising) 33 \rf and 26 \af as the sample
of ``known lenses''. Note that the ``known lens'' sample is a subset of
the ``known lens candidates'' sample.

@@

\subsection{Image presentation in \sw}
\label{sec:data:impres}
In order to perform a blind lens search over the entire \cfhtls
WIDE, we present the volunteers with cutouts of images selected
from the survey region. We briefly describe the image presentation here
for completeness but more information can be found in Paper I. We use
the $g$, $r$ and $i$-band imaging from \cfhtls which are most useful for
visual identification of lenses.  We made color composite images using
the publicly-available code, \humvi\footnote{The open source  color image
composition code used in this work is available from
\texttt{http://github.com/drphilmarshall/HumVI}} following the
prescription of \citet{Lupton2004}. The color scales are chosen to
maximize the contrast between faint extended objects. These parameters
are then fixed during the production of all the tiles, in order to allow
straightforward comparison between one image and another, and for
intuition to be built up about the appearance of stars and galaxies
across the survey.

We extract contiguous cutouts of size 81.84\arcsec\ (440 pixels) with an
overlapping region of 10\arcsec (54 pixels) between the neighbouring
cutouts. This  results in a catalog of $\sim$ 430 000 cutouts for the entire \cfhtls
WIDE region. The size of the individual cutout is determined by
optimising factors such as the typical angular scales of gravitational
lenses, the number of objects seen in a single cutout and the total
number of image cutouts in the survey. If a lens
candidate happens to be too close to the edge of a cutout, then the overlap
between neighbouring cutout allows a volunteer to get a clearer view
of the same candidate in at least one of the cutouts. We note that since
the images are shown randomly, a volunteer may not necessarily come
across the neighbouring cutout unless {\bf they classify} a large
number of images. This is not a problem since our user base is extremely
large and we receive multiple classifications for the same cutout.

@@

The simulated lenses are important to train citizen scientists who are
novice to the task of finding lenses but they are also crucial for
analyzing the classifications performed by the citizen scientists (more
details can be found in Paper I, but see \Sref{sec:swap} for a brief
summary). In this section, we describe the framework used for generating the
simulated lens sample and give details of the sample itself along with
some known limitations.

@@

For the purpose of generating simulated lens systems, we divide them
into two main categories a) galaxy scale lenses b) and group or
cluster-scale lenses. We further subdivide galaxy scale lenses based on the
type of the background sources, namely galaxies and quasars. We do not simulate
group-scale  quasar lenses as they are expected to be even more rare. We now
describe our procedure to generate these different types of lens systems.

@@

Since we want to produce realistic looking lens systems, we simulate
lenses in each of the five \cfhtls~filters. The colors of the background
galaxies are drawn randomly from the photometric CFHTLenS catalog
\citep{Hildebrandt2012,Erben2013}.  Similarly, we use a quasar catalog
from the SDSS Data Release 9 \citep{Paris2012} from which colors are
drawn to simulate quasar lenses. Next, we assumed a Gaussian
profile\footnote{This was due to an oversight. We intend to use either
an exponential or {\bf de Vaucouleurs'} profile.} for
the galaxies. The ellipticity and the position angle (PA) are drawn
randomly from within the range given in Table~\ref{tab:thresh}. The
effective radius of the galaxy is estimated from the luminosity$-$size
relation \citep{Bernardi2003} (with a redshift scaling, to account for
size evolution) given by
\be
R_{\rm eff}= 10^{0.52} \frac{L_r^{2/3}}{{(1+z_s)}^2} \, {\bf Kpc}
\ee
where $L_r=L_s/10^{10.2} {\bf L_\odot}$. On the other hand, quasars are assumed to be
point sources and the PSF, with which quasars are convolved, is assumed
to have a Gaussian profile. The full-width-at-half-maximum of PSF is
equated to that of the median seeing for every filter. The median seeing
values are taken from Table 4 of the official Terapix T0007 release
explanatory document \footnote{
\texttt{http://terapix.iap.fr/cplt/T0007/doc/T0007-doc.pdf}}.

@@

\subsection{Simulated lens sample and catalog description}

@@

We produce catalogs with lens and source properties for each of the
three types of lenses.  The catalogs typically have lens position,
redshift, magnitudes, Einstein radius, ellipticity (whenever available)
and shear (for galaxy-scale lenses only). For the background sources, we
provide the offset from the lens center, redshift, magnitudes, total
magnification, number of lensed images. Additionally, when possible,
ellipticity and effective radius of the background galaxies have also
been provided. These catalogs are available from
\texttt{https://github.com/anupreeta27/SIMCT} and simulated lens sample can be
requested from the authors.

@@

\subsection{Limitations of the simulated sample}
The simulated lens sample, although realistic, is not perfect, due to
the simplicity of the lensing models and our limited understanding of
the uncertainties in the model parameters. Comments from citizen
scientists were very helpful in order to sort some of these failures.
Here, we describe some of the cases or aspects in which the simulations
were known to have failed to look realistic. The fraction of such
simulated sample is small (roughly 5 percent)\footnote{This estimate is
based on the number of \#simfail tags from TALK, the discussion forum.}.

The parameters required by various scaling relations and the models
primarily depend on the photometry of the galaxies, groups and quasars
detected in the survey. For galaxy-scale lenses, the lensing galaxies at
higher redshifts or which are fainter have poor photometric redshift
measurements. Consequently these galaxies are assigned  wrong
luminosity and velocity dispersion estimates. This then results in
simulated lenses which look implausible or unrealistic.  For example,
the lensed images for some of the failed simulations have larger image
separation than what you expect from the luminosity and/or size of the
galaxy. We roughly expect mass to follow light so more massive galaxy
typically looks brighter and/or bigger.

@@

We selected a sample of 450 duds for the \StageOne classification in \sw
and a sample of 500 false positives for the \StageTwo inspection. The
sample of false positives was selected from the candidates which passed
the \StageOne of \sw. We note that this is the first time, we have a
systematically compiled sample of visually inspected false positives by
the \sw volunteers and categorized by the science team. We produce
additional larger sample of a few thousand FPs by scanning through the
low probability images after the completion of \StageTwo. All of these
data products will be made available at
\texttt{http://spacewarps.org/\#/projects/CFHTLS/}. Such a sample has tremendous
utility for training and testing of various lens finding algorithms
\citep[e.g.,][]{Chan2014}.

\begin{figure}
\begin{center}
\includegraphics[scale=0.6]{cand_fp_P_frac.pdf}
\caption{ \label{fig:stackP}
Distribution of different types of candidates as a function of the
posterior probability P. The types of the candidates are the false
positives (FPs), the new candidates and the known candidates. The new
and known candidates have higher detection rate for higher values of P,
as expected.}
\end{center}
\end{figure}

@@

\section{Methodology to produce the \sw-\cfhtls lens sample}
\label{sec:swap}
The \sw works as a single unified system which uses the method of visual
inspection to find gravitational lenses. For the first \sw lens search,
the volunteers were shown images at two stages. At \StageOne, volunteers
are requested to undertake a rapid inspection to select lens candiates
ranging from possible lenses to almost certain lenses. At \StageTwo,
volunteers carefully inspect the candidates from \StageOne and select
only promising lens candidates. A daily snapshot of classifications
performed by volunteers are provided to the science team every night.
This daily batch is analyzed by the Space Warps Analysis Pipeline
(SWAP). The philosophy and the details of SWAP are described in detail
in Paper I, here we briefly summarize how it works.

@@

\subsection{\sw-\cfhtls lens sample}
\label{sec:swlens}

In this section, we describe the \sw-\cfhtls lens sample (henceforth,
called the \sw lens sample). We find a total of 141 candidates with
G$\ge1.3$ (medium-high probability sample) out of which {\bf 59} are new
systems. This sample is further divided as follows. We have a total of
50 candidates with $1.3\le$G$<2$ (medium probability sample) out of
which 30 are new. The quality of candidates in this category is such
that at least one of the inspectors thought the candidate is most
probably a lens (that is, a grade of 2) and a second inspector thought
this is possibly a lens (that is, a grade of 1). Among our high
probability sample (G$\ge$2), there are a total of 91 candidates out of
which {\bf 29} are new. In this category, the minimum grade by all of the
inspectors was 2 suggesting that the candidates are probably or almost
certainly lenses according to all of the inspectors. To avoid
duplication, further information on \sw candidates that were previously
identified in the literature will be made available at
\texttt{http://spacewarps.org/\#/projects/CFHTLS/} whereas the sample of new lens
candidates with G$\ge 1.3$ (medium-high probability) is presented in
this paper (see \Sref{sec:results:newcand}).

We find a total of 288 (and {\bf 245} new) candidates with averaged grade
$0<$G$<1.3$ (low probability sample) which means - in the worst case, at
least one of the inspector thought the candidate is possibly a lens and
in the best case, all three inspectors thought the candidate is possibly
a lens (that is, a grade of 1)\footnote{If grades from the inspectors
were found to be discrepant by 2 or more, these were discussed and
re-graded to resolve the discrepancy.}. Further information on the low
probability sample such as their positions and images will be available
at \texttt{http://spacewarps.org/\#/projects/CFHTLS/}. Note that if all of the
inspectors gave a grade of 0 to a candidate, then it is discarded from
the sample.

@@

We show images of our new sample in \Fref{fig:lc} which is arranged
first in the decreasing order of their grades and then increasing order
of their positions. As the first lens search was a blind search with no
preselection of candidates from any algorithm, we find various types of
lenses, as expected from such a search. The final sample consists of
both galaxy and groups-scale lens candidates. There are detections of
elongated arcs and some interesting point-like quasar lensed images.
Most of them are brighter in bluer $g$ band but some candidates brighter
in the redder $i$ band are also found. Since robotic lens searches often
look for blue colored lensed features, they are very much likely to miss
such interesting lens candidates. We did not find any examples of
exotic lens candidates from the visually inspected P$>0.3$ sample. There
may be some interesting candidates lurking in TALK but these will have
to be sifted through.

@@

The power of citizen scientists also lies in the high dynamic range that
allows us to find systems which vary from short (thick) to long (thin)
arcs, from highly compact to low surface brightness images, from round
and point-like to elongated and curved images, from blue to red, from
regular to exotic kinds of lenses while keeping the false positive rate
low compared to algorithms. Discovery of this large sample of completely
new candidates missed by some of these algorithms demonstrates that the
\sw system is functioning well and the self-taught citizen scientists
have managed to beat the robots in some sense.
